Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1)

### 2014-04-09 WSSSPE1_report/WSSSPE1_summary_report.tex
Adds affiliation, email, and support ack for HL.
Fixes double quote error, and doubly defined label.

\thanks{{}$^{(1)}$ National Science Foundation, Arlington, VA, USA; Computation Institute, University of Chicago \& Argonne National Laboratory, Chicago, IL, USA; \url{d.katz@ieee.org}}
%
\thanks{{}$^{(2)}$ xxx}
%
\thanks{{}$^{(3)}$ Argonne National Laboratory, Chicago, IL, USA; \url{ketan@anl.gov}}
%
\thanks{{}$^{(4)}$ xxx}
%
\thanks{{}$^{(5)}$ xxx}
%
\thanks{{}$^{(7)}$ xxx}
%
\thanks{{}$^{(8)}$ xxx}
%
\thanks{{}$^{(9)}$ xxx}
%
\thanks{{}$^{(10)}$ NORC at the University of Chicago and the Illinois Institute of Technology, Chicago, IL, USA; (\url{sctchoi@uchicago.edu}}
% 
\thanks{{}$^{(11)}$ xxx}
%
\thanks{{}$^{(12)}$ xxx}
%
\thanks{{}$^{(13)}$ xxx}
%
\thanks{{}$^{(14)}$ xxx}

@@

One group of papers presented frameworks that were primarily about
characteristics of software artifacts, connecting with the long
discourse on software quality. This approach is realized in adjectives
that can be applied to pieces of software but might also extend to
describe software projects.  Thus Calero et al.~\cite{Calero_WSSSPE} propose adding elements to the ISO
standards for measuring software quality. Specifically, they propose an additional
dimension of quality they call ``perdurability" or ability to ``last over
time'' made up of three characteristics: reliability, maintainability
and adaptability. This shows some overlap with the framework employed
by Venters et al.~\cite{Venters_WSSSPE} who employ the adjectives of
``extensibility, interoperability, maintainability, portability,
reusability and scalability,'' anticipating the sorts of work that
people would need to do with a software artifact in the future, ``as
stakeholders requirements, technology and environments evolve and
change.'' Venters et al. argue that these choices are really made very
early because they relate to the architecture of the software and
involve trade offs that ought to be analyzed alongside other
attributes of the software artifact. Lenhardt et
al~\cite{Lenhardt_WSSSPE} compare the software lifecycle to the data
lifecycle to argue for the inclusion of metadata throughout a piece of
software's life (discussing, for example, how it has been built and
tested and what ``data in'' and ``data out'' has been considered). In
addition, their analogy suggests that the software lifecycle might add
a phase of ``preservation'' and draw on understandings of what that
involves from studies of data.  In sum, then, these frameworks focused
on what needs to be accomplished to have more sustainable software.

@@

\subsection{Career Tracks for Scientific Software Developers} \note{lead: Daniel S. Katz} \label{sec:defining-sustainability}\label{sec:career-tracks}

@@

Some of the work by Katz was
supported by the National Science Foundation while working at the
Foundation; any opinion, finding, and conclusions or recommendations
expressed in this material are those of the author(s) and do not
necessarily reflect the views of the National Science Foundation.

>>>>

\thanks{{}$^{(1)}$ National Science Foundation, Arlington, VA, USA; Computation Institute, University of Chicago \& Argonne National Laboratory, Chicago, IL, USA; \url{d.katz@ieee.org}}
%
\thanks{{}$^{(2)}$ National Evolutionary Synthesis Center (NESCent),
  Durham, NC, USA; \url{hlapp@nescent.org}}
%
\thanks{{}$^{(3)}$ Argonne National Laboratory, Chicago, IL, USA; \url{ketan@anl.gov}}
%
\thanks{{}$^{(4)}$ xxx}
%
\thanks{{}$^{(5)}$ xxx}
%
\thanks{{}$^{(7)}$ xxx}
%
\thanks{{}$^{(8)}$ xxx}
%
\thanks{{}$^{(9)}$ xxx}
%
\thanks{{}$^{(10)}$ NORC at the University of Chicago and the Illinois Institute of Technology, Chicago, IL, USA; (\url{sctchoi@uchicago.edu}}
% 
\thanks{{}$^{(11)}$ xxx}
%
\thanks{{}$^{(12)}$ xxx}
%
\thanks{{}$^{(13)}$ xxx}
%
\thanks{{}$^{(14)}$ xxx}

@@

One group of papers presented frameworks that were primarily about
characteristics of software artifacts, connecting with the long
discourse on software quality. This approach is realized in adjectives
that can be applied to pieces of software but might also extend to
describe software projects.  Thus Calero et al.~\cite{Calero_WSSSPE} propose adding elements to the ISO
standards for measuring software quality. Specifically, they propose an additional
dimension of quality they call ``perdurability'' or ability to ``last over
time'' made up of three characteristics: reliability, maintainability
and adaptability. This shows some overlap with the framework employed
by Venters et al.~\cite{Venters_WSSSPE} who employ the adjectives of
``extensibility, interoperability, maintainability, portability,
reusability and scalability,'' anticipating the sorts of work that
people would need to do with a software artifact in the future, ``as
stakeholders requirements, technology and environments evolve and
change.'' Venters et al. argue that these choices are really made very
early because they relate to the architecture of the software and
involve trade offs that ought to be analyzed alongside other
attributes of the software artifact. Lenhardt et
al~\cite{Lenhardt_WSSSPE} compare the software lifecycle to the data
lifecycle to argue for the inclusion of metadata throughout a piece of
software's life (discussing, for example, how it has been built and
tested and what ``data in'' and ``data out'' has been considered). In
addition, their analogy suggests that the software lifecycle might add
a phase of ``preservation'' and draw on understandings of what that
involves from studies of data.  In sum, then, these frameworks focused
on what needs to be accomplished to have more sustainable software.

@@

\subsection{Career Tracks for Scientific Software Developers} \note{lead: Daniel S. Katz} \label{sec:career-tracks}

@@

Some of the work by Katz was
supported by the National Science Foundation while working at the
Foundation; any opinion, finding, and conclusions or recommendations
expressed in this material are those of the author(s) and do not
necessarily reflect the views of the National Science Foundation. HL
was supported by the National Evolutionary Synthesis Center (NESCent), 
NSF #EF-0905606.
